---
title: "Nutrition Studies: attitude scales prereg"
author: Derek Powell
date: February 26, 2019
output: 
  html_notebook: 
    code_folding: hide
---

This preregistered study examined the relationsips among fruits and vegetable (FV) consumption and beliefs about the efficacy of FV consumption for improving health, and motivation to consume FVs.

```{r echo=FALSE}
library(tidyverse)
```

## Data import and preprocessing

```{r}
# load (anonymized) data
header <- colnames(read.csv("../data/nutrition+scales+prereg_February+26,+2019_14.51.csv", header = TRUE))
df <- read.csv("../data/nutrition+scales+prereg_February+26,+2019_14.51.csv", skip = 3, header = FALSE, col.names = header)

# do some preliminary tidying
df <- df %>%
  as_tibble() %>%
  select(workerId, StartDate:Finished, sex:ffq_fruit_plums, ffq_time_Page.Submit, e01:surv_pay, LocationLatitude, LocationLongitude) %>%
  rename(Duration = Duration..in.seconds.)

# and trim bad subjects
df <- df %>%
  filter(check_2==2, check_7==7, ffq_veg_check6==6, ffq_fruit_check30==30) %>%
  select(-contains("check"))
```


Experiment was conducted February 26, 2019, recruiting 300 participants with the hope of reaching a final sample size of at least 250. This was successful as we ended up with n = `r nrow(df)` participants. Qualtrics is called "nutrition scales prereg".

FV consumption was measured with 30-item Food Frequency Questionnaire and 2-item FV intake questionnaire.

## Scale reliability

```{r}

rcor <- function(x,y=NULL, digits=3){
  # rounded correlation
  round(cor(x,y),digits)
}

# corr plot code
## ----------------------------------------

reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd)
  cormat <-cormat[hc$order, hc$order]
}

# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}
# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}
  
## ----------------------------------------
df %>%
  # select(e01:e08) %>%
  select(m01:m07r) %>%
  # select(b01:b12r) %>%
  cor() %>%
  reorder_cormat() %>%
  # get_lower_tri() %>%
  reshape2::melt(na.rm=TRUE) %>%
  ggplot(aes(x=Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  geom_tile() +
  geom_text(aes(label=round(value,2)), size=6*.35) +
  scale_fill_distiller(type="div", palette = "RdYlBu", limit=c(-1,1)) +
  # scale_fill_viridis_c(limit = c(-1,1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust=.85)) +
  theme(aspect.ratio = 1) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(fill="corr.", title="motivation")
```

```{r}
df %>%
  select(e01:e08) %>%
  # select(m01:m08) %>%
  # select(b01:b12r) %>%
  cor() %>%
  reorder_cormat() %>%
  # get_lower_tri() %>%
  reshape2::melt(na.rm=TRUE) %>%
  ggplot(aes(x=Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  geom_tile() +
  geom_text(aes(label=round(value,2)), size=6*.35) +
  scale_fill_distiller(type="div", palette = "RdYlBu", limit=c(-1,1)) +
  # scale_fill_viridis_c(limit = c(-1,1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust=.85)) +
  theme(aspect.ratio = 1) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(fill="corr.", title="efficacy")
```

The efficacy scale looks a little two-dimensional here. Items e06r and eo7r form their own little cluster. These are:

- e06r "Whether you eat well or not won’t make much difference if your genetics predispose you to be unhealthy."
- e07r "As you get older, you are likely to develop health problems regardless of how you eat."

It makes sense they would be most strongly correlated with each other, but disappointing they are not very correlated with the rest of the scale.

### Coefficients alpha

```{r}
effect_alpha <- psych::alpha((df %>% select(e01:e08)), check.keys = TRUE)
motive_alpha <- psych::alpha((df %>% select(m01:m07r)), check.keys = TRUE)

effect_alpha$total$raw_alpha
motive_alpha$total$raw_alpha
```

Reliability coefficients looks decent though not amazing. Still somewhat weak for effectiveness. And unfortunately, alpha would go up if e06r and e07r were dropped:

```{r}
effect_alpha
```


## Fruit and vegetable consumption

```{r}
df %>%
  select(workerId, starts_with("ffq_veg"), starts_with("ffq_fruit")) %>%
  gather(food, value, -workerId) %>%
  group_by(food) %>%
  summarize(Mean = mean(value), se = sd(value)/sqrt(n()) ) %>%
  mutate(ll = Mean + se, ul = Mean-se) %>%
  mutate(category = ifelse(grepl("veg", food), "vegetable","fruit")) %>%
  mutate(food = gsub("ffq_veg_","", food)) %>%
  mutate(food = gsub("ffq_fruit_","", food)) %>%
  ggplot(aes(x=reorder(food,-Mean), y=Mean, ymin=ll, ymax=ul)) +
  geom_pointrange() +
  facet_wrap(~category, scales="free_x") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```

FV consumption as assessed by single-item questions asking about average consumption per day.

```{r}
df %>%
  select(workerId, fruitsperday, vegsperday) %>%
  gather(food, value, -workerId) %>% 
  ggplot(aes(x=value)) +
    geom_histogram() +
  facet_wrap(~food) +
  theme_bw()
```


```{r}
df %>%
  select(workerId, 
         fruitsperday, 
         vegsperday, 
         starts_with("ffq_veg"), 
         starts_with("ffq_fruit")
         ) %>%
  gather(food, value, -workerId, -fruitsperday, -vegsperday) %>%
  mutate(category = ifelse(grepl("veg", food), "vegetable","fruit")) %>%
  group_by(workerId, fruitsperday, vegsperday, category) %>%
  summarize(ffq_avg = sum(value)/30) %>%
  spread(category, ffq_avg) -> fv_df

```

Correlation between FFQ and single-item measures are moderately correlated for fruits `r rcor(fv_df$fruitsperday, fv_df$fruit)` and vegetables `r rcor(fv_df$vegsperday, fv_df$vegetable)`. Correlations are similar for both.

As this plot shows, FFQ systematically overestimates the per-day intake compared with the single-item average measurements. Dashed line shows 1:1 relationship.

```{r}
fv_df %>%
  mutate(ffq = fruit + vegetable, single_items = fruitsperday+vegsperday) %>% 
  ggplot(aes(x=single_items, y=ffq)) +
  geom_point(shape=1) +
  # geom_smooth(method="lm") +
  geom_abline(slope=1,intercept=0, linetype="dashed", alpha=.5) +
  xlim(c(0,21)) +
  ylim(c(0,21)) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  labs(title="FV intake: FFQ vs single-items")
```

There appears to be at least one extreme outlier/abberant responder (20 on ffq, 2.5 on si).

## Looking at efficacy and motivation

```{r}
df %>%
  gather(food, value, starts_with("ffq_veg"), starts_with("ffq_fruit")) %>%
  mutate(category = ifelse(grepl("veg", food), "vegetable", "fruit")) %>%
  group_by(workerId, fruitsperday, vegsperday, category) %>%
  summarize(ffq_avg = sum(value) / 30) %>%
  spread(category, ffq_avg) %>%
  left_join(
    (
      df %>%
        gather(item, response, e01:e08, m01:m07r) %>% # exclude b08
        mutate(
          scale = case_when(
            grepl("m", item) ~ "motivation",
            grepl("e", item) ~ "efficacy",
          )
        ) %>%
        group_by(workerId, scale) %>%
        summarize(mean_response = mean(response)) %>%
        spread(scale, mean_response)
    ),
    by = "workerId"
  ) %>%
  mutate(total_si = fruitsperday + vegsperday, total_ffq = fruit + vegetable) %>%
  ungroup() -> df_basic

```

Let's first look at correlations between our three scales and our measures of FV intake. Here's a correlation table:

```{r}

rcor(
  df_basic %>% 
    select(motivation, efficacy, total_si, total_ffq)
  )

```

This is looking pretty good. The correlations between motivaiton and efficacy are quite high, as might be expected. However, some of it is driven by some participants on the extremes.

```{r}
df_basic %>%
  ggplot(aes(x=efficacy, y = motivation)) +
  geom_jitter() +
  # scale_color_viridis_c() +
  geom_smooth(method="lm", color="orange") +
  theme_bw() +
  theme(aspect.ratio=1)
```


Let's press on and look at how motivation and efficacy affect FV intake. Here's a plot of motivaiton and ffq intake. 

```{r}
df_basic %>%
  gather(measure, fv_intake, total_ffq,total_si) %>%
  ggplot(aes(x=motivation, y = fv_intake)) +
  geom_jitter() +
  scale_color_viridis_c() +
  geom_smooth(method="lm", color="orange") +
  theme_bw() +
  facet_wrap(~measure)
```

The relationship between efficacy beliefs and FV consumption is a bit less strong, so we would really have to be looking for our intervention to directly increase both efficacy beliefs and motivaitons.

```{r}
df_basic %>%
  gather(measure, fv_intake, total_ffq,total_si) %>%
  ggplot(aes(x=efficacy, y = fv_intake)) +
  geom_jitter() +
  scale_color_viridis_c() +
  geom_smooth(method="lm", color="orange") +
  theme_bw() +
  facet_wrap(~measure)
```

### Mediation

This is a quick-and-dirty test of mediation. Putting both efficacy and motivation into the same model leads the efficacy coefficient to go to zero (more or less), supporting a mediation model:

`efficacy --> (+) motivation --> (+) FV consumption`

```{r}
df_basic_scaled <-  df_basic %>%
  select(-workerId) %>%
  mutate_all(scale)

lm(total_si ~ efficacy, data=df_basic_scaled) %>% summary()
lm(total_si ~ efficacy + motivation, data=df_basic_scaled) %>% summary()
# lm(total_ffq ~ efficacy + motivation, data=df_basic_scaled) %>% summary()
```

The correlation of motivation with the single-item ffq is r = `r rcor(df_basic$motivation, df_basic$total_si)` and with the full ffq is r = `r rcor(df_basic$motivation, df_basic$total_ffq)`. The first of those especially seems pretty respectable, accounting for `r (cor(df_basic$motivation, df_basic$total_si)^2 %>% round(3))*100`% of the variance.

## Demographics and Possible barriers

What are some possible barriers to FV consumption we might anticipate based on demographics?

### Income

Let's look at participants' income as a predictor of FV intake. We had 6 income brackets:

1 - Less than \$20,000  
2 - \$20,000 - \$30,0000  
3 - \$30,001 - \$50,000  
4 - \$50,001 - \$70,000  
5 - \$70,001 - \$100,000  
6 - More than \$100,000  
0 - Prefer not to say  


```{r}
df_basic %>%
  left_join(df %>% select(workerId, income, educ, age, sex), by ="workerId") %>%
  filter(income!=0) %>%
  ggplot(aes(x=income, y=total_ffq)) +
  geom_jitter(width=.25,height=0) +
  geom_smooth() +
  theme_bw() +
  labs(title="Income x ffq")

```

Not much to see.

### Other demographics

Looking quickly at education and age, nothing pops out.

```{r}
df_basic %>%
  left_join(df %>% select(workerId, income, educ, age, sex), by ="workerId") %>%
  filter(educ!=-1) %>%
  ggplot(aes(x=educ, y=total_ffq)) +
  geom_jitter(height=0,width=.25) +
  geom_smooth() +
  theme_bw() +
  ylim(0,15) +
  labs(title="Education")
```


```{r}
df_basic %>%
  left_join(df %>% select(workerId, income, educ, age, sex), by ="workerId") %>%
  # filter(educ!=-1) %>%
  ggplot(aes(x=age, y=total_ffq)) +
  geom_jitter(height=.1,width=1) +
  geom_smooth() +
  ylim(0,15) +
  theme_bw() +
  labs(title="Age")
```

Not much to see. This is all with ffq measures of fv consumption, things look similar but maybe slightly more interesting for age with single-item measures.

### location?

This is crude, but we can look at how participant's locations relate to their fruit and vegetable intake. Here I will just use latitude as recorded by Qualtrics, on the thinking that southern states have more growing season weather and thus more access to fruits and vegetables. There's a relationship but it's quite weak, and doesn't seem to do anything to account for variance not accounted for by the motivation scale.

```{r}
df_basic %>%
  left_join(df %>% select(workerId, LocationLatitude, LocationLongitude ), by ="workerId") %>%
  filter(LocationLatitude < 50, LocationLatitude > 25) %>%
  ggplot(aes(x=LocationLatitude, y=total_ffq)) +
  # geom_jitter(height=.1,width=1) +
  geom_point() +
  geom_smooth(method="lm") +
  ylim(0,15) +
  theme_bw() +
  labs(title="Latitude x FFQ")


df_basic %>%
  left_join(df %>% select(workerId, LocationLatitude, LocationLongitude ), by ="workerId") %>%
  filter(LocationLatitude < 50, LocationLatitude > 25) %>%
  gather(measure, fv_intake, total_ffq, total_si) %>%
  ggplot(aes(x=motivation, y = fv_intake, color=LocationLatitude)) +
  geom_jitter(width=.1,height=.1) +
  scale_color_viridis_c() +
  geom_smooth(method="lm", color="orange") +
  theme_bw() +
  facet_wrap(~measure)
```

### Miscellany

We could also treat ffq responses as ordinal/likert and then average them, rather than trying to use them to compute daily fruit and vegetable intake. Doing so doesn't meaningfully change the relationships here though. 


```{r}
df %>%
  select(workerId, 
         fruitsperday, 
         vegsperday, 
         starts_with("ffq_veg"), 
         starts_with("ffq_fruit")
         ) %>%
  gather(food, value, -workerId, -fruitsperday, -vegsperday) %>%
  mutate(value = recode( value,
    `1`=1,
    `2.5`=2,
    `6`=3,
    `16`=4,
    `30`=5
  )) %>%
  mutate(category = ifelse(grepl("veg", food), "vegetable","fruit")) %>%
  group_by(workerId, fruitsperday, vegsperday, category) %>%
  summarize(ffq_avg = sum(value)) %>%
  ungroup() %>%
  spread(category, ffq_avg) %>%
  mutate(ffq = fruit + vegetable, single_items = fruitsperday+vegsperday) %>% 
  select(ffq, single_items, workerId) %>%
  left_join(df_basic, by="workerId") %>%
  select(ffq, single_items, motivation, efficacy) %>%
  rcor()
```

# Provisional Conclusions

I plan to conduct a more careful version of these analyses (especially the mediation aspects). But, provisionally I can conclude:

1. Motivation adn efficacy scales are reasonably unidimensional and reliable (though efficacy is not quite unidimensional).
1. Motivation and Efficacy beliefs strongly related
2. Motivation and Efficacy scales robustly predict FV consumption
3. Relationship between efficacy beliefs and FV consumption mediated by Motivation

All in all it's quite a sensible set of results and it suggests that educational interventions that can affect motivation and efficacy beliefs would have promise for improving people's actual fruit and vegetable consumption.

# Updates

March 1, 2019  

After meeting and discussing, the non-unidimensionality of the efficacy scale seems potentially problematic. The easiest solution is to just drop the two items that weren't part of the main cluster (unfortunatley, that leaves us with no reverse coded items).

Let's see what happens if we do that ...

### Coefficient alpha

```{r}
psych::alpha((df %>% select(e01:e05,e08)), check.keys = TRUE)
```

### Correlations and mediation

```{r}
df %>%
  gather(food, value, starts_with("ffq_veg"), starts_with("ffq_fruit")) %>%
  mutate(category = ifelse(grepl("veg", food), "vegetable", "fruit")) %>%
  group_by(workerId, fruitsperday, vegsperday, category) %>%
  summarize(ffq_avg = sum(value) / 30) %>%
  spread(category, ffq_avg) %>%
  left_join(
    (
      df %>%
        gather(item, response, e01:e05, e08, m01:m07r) %>% # exclude b08
        mutate(
          scale = case_when(
            grepl("m", item) ~ "motivation",
            grepl("e", item) ~ "efficacy",
          )
        ) %>%
        group_by(workerId, scale) %>%
        summarize(mean_response = mean(response)) %>%
        spread(scale, mean_response)
    ),
    by = "workerId"
  ) %>%
  mutate(total_si = fruitsperday + vegsperday, total_ffq = fruit + vegetable) %>%
  ungroup() -> df_basic2

rcor(
  df_basic2 %>% 
    select(motivation, efficacy, total_si, total_ffq)
  )

df_basic_scaled2 <-  df_basic2 %>%
  select(-workerId) %>%
  mutate_all(scale)

lm(total_si ~ efficacy, data=df_basic_scaled2) %>% summary()
lm(total_si ~ efficacy + motivation, data=df_basic_scaled2) %>% summary()
```

All correlations and the mediation relationships look essentially the same with the new scale.
