---
title: "Nutrition Studdies: barriers pilot"
output: 
  html_notebook: 
    code_folding: hide
---

This experiment examined the relationsips among fruits and vegetable (FV) consumption and beliefs about the efficacy of FV consumption for improving health, motivation to consume FVs, and of experienced barriers to FV consumption.

Experiment was conducted February 12, 2019, recruiting 220 participants. Qualtrics is called "nutrition barriers pilot".

This is a pilot study conducted to test a new scale measuring barriers to fruit and vegetable consumption. The idea is that accounting for these barriers may reveal a stronger relationship between motivation and efficacy beliefs and FV consumption. If this suggestion is validated, it would further validate those scales as meaningfully connected to FV consumption. Further, to the extent that an intervention affects responses on these scales, we can conclude it is affecting beliefs that are meaningfully related to FV consumption.

FV consumption was measured with 30-item Food Frequency Questionnaire and 2-item FV intake questionnaire.

```{r echo=FALSE}
library(tidyverse)
```

## Data import and preprocessing

```{r}
# load (anonymized) data
header <- colnames(read.csv("../data/nutrition+barriers+pilot_February+12,+2019_14.46.csv", header = TRUE))
df <- read.csv("../data/nutrition+barriers+pilot_February+12,+2019_14.46.csv", skip = 3, header = FALSE, col.names = header)

# do some preliminary tidying
df <- df %>%
  as_tibble() %>%
  select(workerId, StartDate:Finished, sex:ffq_fruit_plums, ffq_time_Page.Submit, e01:surv_pay, LocationLatitude, LocationLongitude) %>%
  rename(Duration = Duration..in.seconds.) %>%
  mutate_at(vars(m08, b01,b03,b04, b06,b08,b09,b10,b11), function(x){8 - x}) # fixing issues in qualtrics

# and trim bad subjects
df <- df %>%
  filter(check_2==2, check_7==7, ffq_veg_check6==6, ffq_fruit_check30==30) %>%
  select(-contains("check"))
```


## Scale reliability

```{r}

rcor <- function(x,y=NULL, digits=3){
  # rounded correlation
  round(cor(x,y),digits)
}

# corr plot code
## ----------------------------------------

reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd)
  cormat <-cormat[hc$order, hc$order]
}

# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}
# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}
  
## ----------------------------------------
df %>%
  # select(e01:e08) %>%
  select(m01:m08) %>%
  # select(b01:b12r) %>%
  cor() %>%
  reorder_cormat() %>%
  # get_lower_tri() %>%
  reshape2::melt(na.rm=TRUE) %>%
  ggplot(aes(x=Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  geom_tile() +
  geom_text(aes(label=round(value,2)), size=6*.35) +
  scale_fill_distiller(type="div", palette = "RdYlBu", limit=c(-1,1)) +
  # scale_fill_viridis_c(limit = c(-1,1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust=.85)) +
  theme(aspect.ratio = 1) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(fill="corr.", title="motivation")
```

```{r}
df %>%
  select(e01:e08) %>%
  # select(m01:m08) %>%
  # select(b01:b12r) %>%
  cor() %>%
  reorder_cormat() %>%
  # get_lower_tri() %>%
  reshape2::melt(na.rm=TRUE) %>%
  ggplot(aes(x=Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  geom_tile() +
  geom_text(aes(label=round(value,2)), size=6*.35) +
  scale_fill_distiller(type="div", palette = "RdYlBu", limit=c(-1,1)) +
  # scale_fill_viridis_c(limit = c(-1,1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust=.85)) +
  theme(aspect.ratio = 1) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(fill="corr.", title="efficacy")
```


```{r}
df %>%
  # select(e01:e08) %>%
  # select(m01:m08) %>%
  select(b01:b12r) %>%
  cor() %>%
  reorder_cormat() %>%
  # get_lower_tri() %>%
  reshape2::melt(na.rm=TRUE) %>%
  ggplot(aes(x=Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  geom_tile() +
  geom_text(aes(label=round(value,2)), size=6*.35) +
  scale_fill_distiller(type="div", palette = "RdYlBu", limit=c(-1,1)) +
  # scale_fill_viridis_c(limit = c(-1,1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust=.85)) +
  theme(aspect.ratio = 1) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(fill="corr.", title="Barriers")
```

So a first pass here looks like b08 is a problem item, being signifiantly negatively correlated with some other things. This item measured concerns about pesticides (I suppose only health nuts worry about that?).

I'm going to augment the scale by removing b08 (though this doesn't end up making much difference).

```{r}
effect_alpha <- psych::alpha((df %>% select(e01:e08)), check.keys = TRUE)
motive_alpha <- psych::alpha((df %>% select(m01:m08)), check.keys = TRUE)
barriers_alpha <- psych::alpha((df %>% select(b01:b07r,b09:b12r)))

effect_alpha$total$raw_alpha
motive_alpha$total$raw_alpha
barriers_alpha$total$raw_alpha
```

Reliability coefficients looks decent though not amazing. Still somewhat weak for effectiveness.

## Fruit and vegetable consumption

```{r}
df %>%
  select(workerId, starts_with("ffq_veg"), starts_with("ffq_fruit")) %>%
  gather(food, value, -workerId) %>%
  group_by(food) %>%
  summarize(Mean = mean(value), se = sd(value)/sqrt(n()) ) %>%
  mutate(ll = Mean + se, ul = Mean-se) %>%
  mutate(category = ifelse(grepl("veg", food), "vegetable","fruit")) %>%
  mutate(food = gsub("ffq_veg_","", food)) %>%
  mutate(food = gsub("ffq_fruit_","", food)) %>%
  ggplot(aes(x=reorder(food,-Mean), y=Mean, ymin=ll, ymax=ul)) +
  geom_pointrange() +
  facet_wrap(~category, scales="free_x") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```

FV consumption as assessed by single-item questions asking about average consumption per day.

```{r}
df %>%
  select(workerId, fruitsperday, vegsperday) %>%
  gather(food, value, -workerId) %>% 
  ggplot(aes(x=value)) +
    geom_histogram() +
  facet_wrap(~food) +
  theme_bw()
```


```{r}
df %>%
  select(workerId, 
         fruitsperday, 
         vegsperday, 
         starts_with("ffq_veg"), 
         starts_with("ffq_fruit")
         ) %>%
  gather(food, value, -workerId, -fruitsperday, -vegsperday) %>%
  mutate(category = ifelse(grepl("veg", food), "vegetable","fruit")) %>%
  group_by(workerId, fruitsperday, vegsperday, category) %>%
  summarize(ffq_avg = sum(value)/30) %>%
  spread(category, ffq_avg) -> fv_df

```

Correlation between FFQ and single-item measures are moderately correlated for fruits `r rcor(fv_df$fruitsperday, fv_df$fruit)` and vegetables `r rcor(fv_df$vegsperday, fv_df$vegetable)`. Correlations are similar for both.

As this plot shows, FFQ systematically overestimates the per-day intake compared with the single-item average measurements. Dashed line shows 1:1 relationship.

```{r}
fv_df %>%
  mutate(ffq = fruit + vegetable, single_items = fruitsperday+vegsperday) %>% 
  ggplot(aes(x=single_items, y=ffq)) +
  geom_point() +
  # geom_smooth(method="lm") +
  geom_abline(slope=1,intercept=0, linetype="dashed", alpha=.5) +
  # xlim(c(0,14)) +
  # ylim(c(0,14)) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  labs(title="FV intake: FFQ vs single-items")
```


**February 16, 2019 (AFTER MEETING)**: Can also treat ffq responses as ordinal/likert and then average. Doing so do dosn't improve correlations though ...


```{r}
df %>%
  select(workerId, 
         fruitsperday, 
         vegsperday, 
         starts_with("ffq_veg"), 
         starts_with("ffq_fruit")
         ) %>%
  gather(food, value, -workerId, -fruitsperday, -vegsperday) %>%
  mutate(value = recode( value,
    `1`=1,
    `2.5`=2,
    `6`=3,
    `16`=4,
    `30`=5
  )) %>%
  mutate(category = ifelse(grepl("veg", food), "vegetable","fruit")) %>%
  group_by(workerId, fruitsperday, vegsperday, category) %>%
  summarize(ffq_avg = sum(value)) %>%
  ungroup() %>%
  spread(category, ffq_avg) %>%
  mutate(ffq = fruit + vegetable, single_items = fruitsperday+vegsperday) %>% 
  select(ffq, single_items) %>%
  rcor()
```

## Looking at efficacy and motivation

```{r}
df %>%
  gather(food, value, starts_with("ffq_veg"), starts_with("ffq_fruit")) %>%
  mutate(category = ifelse(grepl("veg", food), "vegetable", "fruit")) %>%
  group_by(workerId, fruitsperday, vegsperday, category) %>%
  summarize(ffq_avg = sum(value) / 30) %>%
  spread(category, ffq_avg) %>%
  left_join(
    (
      df %>%
        gather(item, response, e01:e08, m01:m08, b01:b07r, b09:b12r) %>% # exclude b08
        mutate(
          scale = case_when(
            grepl("m", item) ~ "motivation",
            grepl("e", item) ~ "efficacy",
            grepl("b", item) ~ "barriers",
          )
        ) %>%
        group_by(workerId, scale) %>%
        summarize(mean_response = mean(response)) %>%
        spread(scale, mean_response)
    ),
    by = "workerId"
  ) %>%
  mutate(total_si = fruitsperday + vegsperday, total_ffq = fruit + vegetable) %>%
  ungroup() -> df_basic

```


Let's first look at correlations between our three scales and our measures of FV intake. Here's a correlation table:

```{r}

rcor(
  df_basic %>% 
    select(barriers, motivation, efficacy, total_si, total_ffq)
  )

```

Some disappointing things here. First, barriers isn't meaningfully correlated with our ffq-based measure of fv intake, and only weakly correlated with our single-item measures. Worse, barriers is negatively correlated with motivation and efficacy.


Let's press on and look at how motivation and efficacy affect FV intake. Here's a plot of motivaiton and ffq intake. 

```{r}
df_basic %>%
  gather(measure, fv_intake, total_ffq,total_si) %>%
  ggplot(aes(x=motivation, y = fv_intake)) +
  geom_jitter() +
  scale_color_viridis_c() +
  geom_smooth(method="lm", color="orange") +
  theme_bw() +
  facet_wrap(~measure)
```

The relationship between efficacy beliefs and FV consumption is a bit less strong, so we would really have to be looking for our intervention to directly increase both efficacy beliefs and motivaitons.

```{r}
df_basic %>%
  gather(measure, fv_intake, total_ffq,total_si) %>%
  ggplot(aes(x=efficacy, y = fv_intake)) +
  geom_jitter() +
  scale_color_viridis_c() +
  geom_smooth(method="lm", color="orange") +
  theme_bw() +
  facet_wrap(~measure)
```

Putting both efficacy and motivation into the same model leads the efficacy coefficient to go to zero (more or less), supporting a mediation model:

`efficacy --> (+) motivation --> (+) FV consumption`

```{r}
df_basic_scaled <-  df_basic %>%
  select(-workerId) %>%
  mutate_all(scale)

lm(total_si ~ efficacy + motivation, data=df_basic_scaled) %>% summary()
lm(total_ffq ~ efficacy + motivation, data=df_basic_scaled) %>% summary()
```

The correlation of motivation with the single-item ffq is r = `r rcor(df_basic$motivation, df_basic$total_si)` and with the full ffq is r = `r rcor(df_basic$motivation, df_basic$total_ffq)`. The first of those especially seems pretty respectable, accounting for `r cor(df_basic$motivation, df_basic$total_si)^2 %>% round(3)`% of the variance.

### Barriers

Now let's add barriers to the mix. We can plot how barriers might affect the relationship between fv intake and motivation, by adding color to our earlier plots.

```{r}
df_basic %>%
  gather(measure, fv_intake, total_ffq, total_si) %>%
  ggplot(aes(x=motivation, y = fv_intake, color=barriers)) +
  geom_jitter() +
  scale_color_viridis_c() +
  geom_smooth(method="lm", color="orange") +
  theme_bw() +
  facet_wrap(~measure)
```

This isn't showing what I might have expected, which would have been people with higher barriers (yellow dots) being below the regression line and people with low barriers (blue dots) above the regression line. This almost looks like the opposite, to the extent it looks like anything at all.

One clear issue is that barriers is pretty strongly correlated with motivation and efficacy. 

```{r}
df_basic %>%
  ggplot(aes(x=barriers, y = motivation)) +
  geom_point() +
  geom_smooth(method="lm", color="orange") +
  theme_bw() +
  theme(aspect.ratio=1)

df_basic %>%
  ggplot(aes(x=barriers, y = efficacy)) +
  geom_point() +
  geom_smooth(method="lm", color="orange") +
  theme_bw() +
  theme(aspect.ratio=1)
```

We can also look at the relationship between barriers and FV intake. As in the correlation table, very little correlation.

```{r}
df_basic %>%
  gather(measure, fv_intake, total_ffq,total_si) %>%
  ggplot(aes(x=barriers, y = fv_intake)) +
  geom_jitter() +
  scale_color_viridis_c() +
  geom_smooth(method="lm", color="orange") +
  theme_bw() +
  facet_wrap(~measure)
```

Yet throwing these guys into a regression together produces a significant coefficient for barriers. So there's some kind of weird suppression thing going on? Adding motivation and efficacy to the same model reveals no effect of efficacy, suggesting a mediated relationship (which is quite sensible). 

So more to unravel here ...

```{r}


lm(total_ffq ~ efficacy + barriers, data=df_basic_scaled) %>% summary()
lm(total_ffq ~ motivation + barriers, data=df_basic_scaled) %>% summary()
lm(total_ffq ~ efficacy + motivation + barriers, data=df_basic_scaled) %>% summary()

lm(total_ffq ~ barriers, data=df_basic_scaled) %>% summary()
```

Controlling for motivaiton, barriers actually has a POSITIVE relationship with FV intake when they are put into the same model. So the barriers scale is pretty clearly not workign the way we had in mind.

## Rethinking barriers

Let's look at the barriers more specifically ....

```{r}
df %>%
  select(workerId, starts_with("ffq_veg"), starts_with("ffq_fruit")) %>%
  gather(food, resp, -workerId) %>%
  group_by(workerId) %>%
  summarize(fv_intake = sum(resp)/30) %>%
  left_join(
    df %>%
      select(workerId, b01:b12r),
    by="workerId"
  ) %>%
  select(-workerId) %>%
  rcor()
```

Looking at the individual barriers, some correlations with FV intake are going the wrong way! Most notably, concerns about pesticides (I suppose only health nuts worry about that?--this problem was already evident before), being disappointed about fruit not being ripe (necessitates purchasing it), fruits being out of season (again, have to know something and pursue fvs), and eating out at places that don't serve FVs (suggests awarness of FVs absence?).

### Income

Let's look at participants' income as a predictor of FV intake and barriers, etc. We had 6 income brackets:

1 - Less than \$20,000  
2 - \$20,000 - \$30,0000  
3 - \$30,001 - \$50,000  
4 - \$50,001 - \$70,000  
5 - \$70,001 - \$100,000  
6 - More than \$100,000  
0 - Prefer not to say  


```{r}
df_basic %>%
  left_join(df %>% select(workerId, income, educ, age, sex), by ="workerId") %>%
  filter(income!=0) %>%
  ggplot(aes(x=income, y=total_ffq)) +
  geom_jitter(width=.25,height=0) +
  geom_smooth() +
  theme_bw() +
  labs(title="Income x ffq")

df_basic %>%
  left_join(df %>% select(workerId, income, educ, age, sex), by ="workerId") %>%
  filter(income!=0) %>%
  ggplot(aes(x=income, y=barriers)) +
  geom_jitter(width=.25,height=0) +
  geom_smooth() +
  theme_bw() +
  labs(title="Income x barriers")

```

Well that's a big nothing! Not much to see, though a bit of a decrease in barriers at high income brackets, presumably due to the items asking about financial hardship.

### Other demographics

Looking quickly at education and age, nothing pops out.

```{r}
df_basic %>%
  left_join(df %>% select(workerId, income, educ, age, sex), by ="workerId") %>%
  filter(educ!=-1) %>%
  ggplot(aes(x=educ, y=total_ffq)) +
  geom_jitter(height=0,width=.25) +
  geom_smooth() +
  theme_bw() +
  ylim(0,15) +
  labs(title="Education")
```


```{r}
df_basic %>%
  left_join(df %>% select(workerId, income, educ, age, sex), by ="workerId") %>%
  # filter(educ!=-1) %>%
  ggplot(aes(x=age, y=total_ffq)) +
  geom_jitter(height=.1,width=1) +
  geom_smooth() +
  ylim(0,15) +
  theme_bw() +
  labs(title="Age")
```

Not much to see. This is all with ffq measures of fv consumption, things look similar but maybe slightly more interesting for age with single-item measures.

### location?

This is crude, but just using latitude, on hte logic that southern states have more growing season weather and thus more access to fruits and vegetables. There's a relationship but it's quite weak, and doesn't seem to do anything to account for variance not accounted for by the motivation scale.

```{r}
df_basic %>%
  left_join(df %>% select(workerId, LocationLatitude, LocationLongitude ), by ="workerId") %>%
  filter(LocationLatitude < 50, LocationLatitude > 25) %>%
  ggplot(aes(x=LocationLatitude, y=total_ffq)) +
  # geom_jitter(height=.1,width=1) +
  geom_point() +
  geom_smooth() +
  ylim(0,15) +
  theme_bw() +
  labs(title="Latitude x FFQ")


df_basic %>%
  left_join(df %>% select(workerId, LocationLatitude, LocationLongitude ), by ="workerId") %>%
  filter(LocationLatitude < 50, LocationLatitude > 25) %>%
  gather(measure, fv_intake, total_ffq, total_si) %>%
  ggplot(aes(x=motivation, y = fv_intake, color=LocationLatitude)) +
  geom_jitter(width=.1,height=.1) +
  scale_color_viridis_c() +
  geom_smooth(method="lm", color="orange") +
  theme_bw() +
  facet_wrap(~measure)
```

Playing around ... nothing to see here.

```{r}

df_aug <- df_basic %>%
  left_join(df %>% select(workerId, income, LocationLatitude, LocationLongitude ), by ="workerId") %>%
  filter(LocationLatitude < 50, LocationLatitude > 25) %>%
  filter(income!=0)

lm(total_ffq ~ motivation + LocationLatitude, data = df_aug) %>% summary()

```

