---
title: "open-ended reliability coding 2"
output: html_notebook
---

This is an analysis of reliability with two new coders, Aarthi and Kristine, and our modestly revised coding system as of July 11, 2019.



```{r}
library(tidyverse)
library(rel)
source("helpers.R")

df_all <- read_qualtrics("../data_private/interv-free-responses-may2019 - reliability2.csv")
df <- df_all[1:52,]
```


```{r}
df <- df %>%
  mutate(
    junk_phyto_d = recode(junk_phyto_d, "?" = .5),
    junk_phyto = recode(junk_phyto, "?" = .5),
    junk_free_radicals_d = recode(junk_free_radicals_d, "?" = .5),
    junk_free_radicals = recode(junk_free_radicals, "?" = .5)
    ) %>%
  mutate_at(vars(contains("junk_")), function(x){ifelse(is.na(x),0,x)}) %>%
  mutate(junk = junk_phyto + junk_phyto_d + junk_free_radicals + junk_free_radicals_d)
```


### Percent Agreement

```{r}
df %>%
  filter(junk==0) %>%
  select(-contains("junk")) %>%
  gather(
    item, code, -workerId
  ) %>%
  mutate(coder = if_else(grepl("_d", item),"Derek","Interns")) %>%
  mutate(item = gsub("_d","",item)) %>%
  spread(coder, code) %>%
  mutate(match = ifelse(Derek==Interns,1,0)) %>%
  group_by(item) %>%
  summarize(agreement = mean(match, na.rm=TRUE))
```


### Cohen's Kappa

Let's compute a real reliability coefficient: the quadratic weighted kappa for each scale.

Some sources on Kappa:

* A [peer reviewed article](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3900052/) reviewing its use
* The [Wikipedia page](https://en.wikipedia.org/wiki/Cohen%27s_kappa)
* A nice [stackoverflow answer](https://stats.stackexchange.com/questions/82162/cohens-kappa-in-plain-english)

```{r}
kappa_weighting <- "quadratic"
df_nojunk <- df %>%
  filter(junk==0)
summary(ckap(select(df_nojunk, phyto, phyto_d), weight = kappa_weighting))
summary(ckap(select(df_nojunk, free_radicals, free_radicals_d), weight = kappa_weighting))
```

## recodes

recoding with kristine after poor reliability on "phyto" question in first round.

```{r}
df2  <- df_all[52:nrow(df_all),]

df2 <- df2 %>%
  select(workerId, phyto, phyto_d, junk_phyto, junk_phyto_d) %>%
    mutate(
    junk_phyto_d = recode(junk_phyto_d, "?" = .5),
    junk_phyto = recode(junk_phyto, "?" = .5),
    # junk_free_radicals_d = recode(junk_free_radicals_d, "?" = .5),
    # junk_free_radicals = recode(junk_free_radicals, "?" = .5)
    ) %>%
  mutate_at(vars(contains("junk_")), function(x){ifelse(is.na(x),0,x)}) %>%
  mutate(junk = junk_phyto + junk_phyto_d)
```

### percent agreement


```{r}
df2 %>%
  filter(junk==0) %>%
  select(-contains("junk")) %>%
  gather(
    item, code, -workerId
  ) %>%
  mutate(coder = if_else(grepl("_d", item),"Derek","Interns")) %>%
  mutate(item = gsub("_d","",item)) %>%
  spread(coder, code) %>%
  mutate(match = ifelse(Derek==Interns,1,0)) %>%
  group_by(item) %>%
  summarize(agreement = mean(match, na.rm=TRUE))
```

```{r}
kappa_weighting <- "quadratic"
df2_nojunk <- df2 %>%
  filter(junk==0)
summary(ckap(select(df2_nojunk, phyto, phyto_d), weight = kappa_weighting))
```

looks good now!

# leafy v carrots

codes from aarthi and kristine, checking reliability against one another. reliability is poor for codes as of 7/26/19, 5:23 PM

```{r}
# from google sheet: leafy_carrots_rel
df3 <- read_tsv(clipboard())

summary(ckap(df3), weight="quadratic")
```

```{r}
df3 %>%
  mutate(match = ifelse(aarthi==kristine,1,0)) %>%
  summarize(agreement = mean(match, na.rm=TRUE))
```

